---
title: "Bayesian Chi Squared Procedures"
author: "Alisa Krasilnikov"
format: html
embed-resources: true
editor: source
---

# References: 
 
1. Doing Bayesian data analysis:
 - Chapter 24 - Count Predicted Variables
 - Exercise 24.3 

2. https://www.flutterbys.com.au/stats/tut/tut11.2b.html


## Data

```{r, message = FALSE, echo = TRUE}
#| label: load-packages
library(brms)
library(tidybayes)
library(dplyr)
```

Our dataset is the Indian food data set from Kaggle <https://www.kaggle.com/datasets/kritirathi/indian-food-dataset-with>. This is a set which pulls data from online Indian food recipes, and classifies them in various categories. We are particularly interested in whether there is a difference in flavor profile (spicy, sweet, bitter, and sour) across diets (vegetarian and non-vegetarian). 

```{r}
#| label: read-csv
indfood <- read.csv(here::here("Ifood_new.csv"))
```

```{r}
#| label: data-cleaning
indfood <- indfood |> 
  filter(flavor_profile != "-1") |>
  mutate(dum_diet = if_else(diet == "vegetarian", 1, 0))

```

```{r}
#| label: summary
summary(indfood)
```

## Bayesian model

### Choosing Priors

We’ll assume the betas and sigma are independent. We believe that if a dish is sweet rather than bitter, then it will most likely be a vegetarian dish, since we couldn't think of many examples of desserts which were not vegetarian. Therefore, we specify an informative prior for the coefficient of the sweet flavor profile that reflects our belief that it increases the probability a dish is vegetarian (i.e., that it pushes the log-odds in a positive direction):

$\beta_{sweet}$ ~ N(0.75, 0.25)

This prior centers the log-odds increase at 0.75 (approximately corresponding to a 68% chance of being vegetarian, all else equal), while still allowing moderate uncertainty. For the other coefficients, we do not have strong prior opinions, and we allow brms to apply its default weakly informative priors.


**Prior predictive dist for sweet dish**
```{r}
n_rep = 10000

beta0 <- rnorm(n_rep, 0, 1)
beta_sweet <- rnorm(n_rep, 0.75, 0.25)

log_odds <- beta0 + beta_sweet * sweet

p <- 1 / (1 + exp(-log_odds))

y_sim <- rbinom(n_rep, size = 1, prob = p)

hist(p,
     xlab = "Prior predicted P(Vegetarian) for sweet dish",
     breaks = 100,
     col = "pink",
     main = "Prior Predictive Distribution")
```
This isn't exactly what we want. The range is a little bit too big, but we like where it's centered. Let's adjust it a little bit so the effect of sweetness is a little bit stronger.

$\beta_{sweet}$ ~ N(1, 0.1)

```{r}
beta_0 <- rnorm(n_rep, 0, 1)
beta_sweet <- rnorm(n_rep, 1, 0.1)

log_odds <- beta0 + beta_sweet

p <- 1 / (1 + exp(-log_odds))

y_sim <- rbinom(n_rep, size = 1, prob = p)

hist(p,
     xlab = "Prior predicted P(Vegetarian) for sweet dish",
     breaks = 100,
     col = "pink",
     main = "Prior Predictive Distribution")
```
This looks much better.

```{r}
n_rep = 1000

# x is binary: 0 (not sweet), 1 (sweet)
x <- sample(c(0, 1), n_rep, replace = TRUE)

# Priors for coefficients
beta0 <- rnorm(n_rep, 0, 1)           # intercept
beta1 <- rnorm(n_rep, 1, 0.1)      # effect for sweet = 1


# Compute probabilities for x = 0 and x = 1
p0 <- plogis(beta0 + beta1 * 0)  # bitter
p1 <- plogis(beta0 + beta1 * 1)  # sweet

# Plot histograms
hist(p0, breaks=100, col=rgb(0,0,1,0.4), xlim=c(0,1), ylim = c(0, 40),
     main="Prior Predictive Distribution of P(Vegetarian)",
     xlab="Probability", ylab="Frequency")
hist(p1, breaks=100, col=rgb(1,0,0,0.4), add=TRUE)
legend("topleft", legend=c("Bitter (Reference)", "Sweet"),
       fill=c(rgb(0,0,1,0.4), rgb(1,0,0,0.4)))

```

### Fitting with brms

```{r}
fit <- brm(
  data = indfood,
  family = bernoulli(link = "logit"),
  dum_diet ~ 1 + flavor_profile,
  prior = c(
    prior(normal(1, 0.1), class = "b", coef = "flavor_profilesweet")
    # No need to set priors for other coefficients, since we're happy with defaults
  ),
  iter = 6000,
  warmup = 1000,
  control = list(max_treedepth = 15),
  chains = 4,
  refresh = 0
)
```

```{r}
summary(fit)
```






